---
title: "code for ENVS193DS HW4"
author: Maddie Manzagol 
data: May 20, 2023
format: docx
---
#Question: How does fish length predict fish weight for trout perch (across all sample years)? 
```{r loading-in-packages}
library(tidyverse)
library(here)
library(lterdatasampler)
library(naniar)
```

```{r set-up-data-using-here-package}
fishdata <- read_csv(here::here("data", "ntl6_v12.csv"))
```

#Problem 1
1. Null and alternative hypotheses in mathematical and biological terms

Mathematical: HO: μ = 0; HA: μ ≠ 0 

Biological: The null hypothesis will be "fish length does not predict fish weight for trout perch"; meanwhile, the alternative hypothesis will be "fish length does predict fish weight for trout perch"

2. Create a visualization of the missing data for the filtered data set containing the observations you will use and also write an accompanying caption explaining how/if the missing data is relevant to your hypotheses. 

```{r filtering-data-set}
fish_summary <- subset(fishdata, select = c(length, weight))
```

```{r visualize-missing-data}
gg_miss_var(fish_summary)
```

The missing data shown from the gg_miss_var plot highlights a substantial amount of missing data for fish weight. This will be important to note because our statistical question is asking if fish length predicts fish weight; however, if a lot of data is missing for fish weight then it will likely be difficult to accurately determine a relationship between fish length and fish weight.

3. Run your test (Linear regression test)
```{r}
fishlinearregression <- lm(weight~length, data= fish_summary)
```

