---
title: "code for ENVS193DS HW4"
author: Maddie Manzagol 
data: May 20, 2023
format: docx
---
#Question: How does fish length predict fish weight for trout perch (across all sample years)? 
```{r loading-in-packages}
library(tidyverse)
library(here)
library(lterdatasampler)
library(naniar)
library(performance)
library(broom)
library(flextable)
library(ggeffects)
library(car)
```

```{r set-up-data-using-here-package}
fishdata <- read_csv(here::here("data", "ntl6_v12.csv"))
```

#Problem 1
1. Null and alternative hypotheses in mathematical and biological terms

Mathematical: HO: Î² = 0; HA: Î² â‰  0 

Biological: The null hypothesis will be "fish length does not predict fish weight for trout perch"; meanwhile, the alternative hypothesis will be "fish length does predict fish weight for trout perch"

2. Create a visualization of the missing data for the filtered data set containing the observations you will use and also write an accompanying caption explaining how/if the missing data is relevant to your hypotheses. 

```{r filtering-data-set}
fish_summary <- subset(fishdata, select = c(length, weight))
```

```{r visualize-missing-data}
gg_miss_var(fish_summary)
```

The missing data shown from the gg_miss_var plot highlights a substantial amount of missing data for fish weight. This will be important to note because our statistical question is asking if fish length predicts fish weight; however, if a lot of data is missing for fish weight then it will likely be difficult to accurately determine a relationship between fish length and fish weight.

3. Run your test (Linear regression test)
```{r}
fishlinearregression <- lm(weight~length, data= fish_summary)

# makes the viewer pane show a 2x2 grid of plots
# format: par(mfrow = c(number of rows, number of columns))
par(mfrow = c(2, 2))
plot(fishlinearregression)
```

plot 1--Residual vs Fitted: The residual vs fitted plot generated from "fishlinearregression" shows residuals that are not evenly and randomly distributed around the horizontal dotted line. Therefore, the assumption of homoscedasticity is violated. 

plot 2--Normal QQ: The normal QQ plot shows residuals that are not entirely linearly distributed, meaning the points do not follow a straight line. There is a portion distributed between quantiles -2 to 2 that follow a very linear line but after quantile 2 the slope changes. 

plot 3--Scale-Location: The scale-location plot highlights that the square root of residuals does not highlight an even and random spread around the red line. This means that the variances change and homoscedasticity is violated. 

plot 4--Residual vs Leverage: The residual vs leverage plot highlights some outliers (labeled) indicating that there are data points that are influencing the model. This means that if the model was calculated without these labeled outliers, then the model would predict a different estimate. 


6.
```{r results-summary}
model_summary <- summary(fishlinearregression)

model_summary
```

7. 
```{r Anova-summary}
model_squares <- anova(fishlinearregression)

model_squares
```


```{r Anova-summary-table}
model_squares_table <- tidy(model_squares) %>% 
  mutate(across(sumsq:meansq,~round(.x, digits = 5))) %>% 
  mutate(statistic = round(statistic, digits = 1)) %>% 
  mutate(p.value = case_when(p.value < 0.001~"<0.001")) %>% 
  mutate(term = case_when(term == "stem_length" ~ "Stem length (mm)",
    TRUE ~ term
  )) %>% 
  # make the data frame a flextable object
  flextable() %>% 
  # change the header labels to be meaningful
  set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of squares",
                    meansq = "Mean squares",
                    statistic = "F-statistic",
                    p.value = "p-value")

model_squares_table
```

8. In 1-2 sentences, describe how the ANOVA table relates to the information you get from the summary() object. 

The ANOVA table relates to the information you get from the summary object because both report f-statistic, the degrees of freedom, and the p-value. 

9. In 2-3 sentences, summarize your results in prose with in-text references to test results. Include all relevant information.

Given that we hypothesized that fish length would predict fish weights, our null hypothesis was that fish length did not predict fish weight. Fish weight was significantly predicted by differing fish lengths (analysis of variance, F(1, 91,088) = 136,700.2, p < 0.001,ðžª= 0.05).



10. Create a visualization with model predictions and confidence intervals on top of the underlying data. Finalize your plot.
a. Write an accompanying caption

```{r}
# extract model predictions using ggpredict
predictions <- ggpredict(fishlinearregression, terms = "length")

predictions

plot_predictions <- ggplot(data = fish_summary, 
                           aes(x = length, y = weight)) +
  # first plot the underlying data from fish_summary
  geom_point() +
  # then plot the predictions
  geom_line(data = predictions, 
            aes(x = x, y = predicted), 
            color = "blue", linewidth = 1) +
  # then plot the 95% confidence interval from ggpredict
  geom_ribbon(data = predictions, 
              aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  # theme and meaningful labels
  theme_bw() +
  labs(x = "Fish Length",
       y = "Fish Weight")

plot_predictions
```

a. The figure above plot fish weight prediction as a function of fish length using data collected about perch trout located in the North Temperate Lakes from 1981 to present day. The data points represent the predicted fish weight given a fish length; meanwhile, the blue line indicates the overall trendline for the predicted correlation between fish length and fish weight. 



